---
title: "p8105_hw3_dd2948"
author: "David DeStephano"
date: "October 7, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(tidyverse)
```

##Problem 1
```{r}
library(p8105.datasets)
data("instacart")

instacart %>%
  group_by(aisle) %>%
  summarize(n = n()) %>% 
  arrange(desc(n))

#Alternatively:
count(instacart, aisle)
#Fresh vegatables is the most popular

```

The dimensions of this data set are `r dim(instacart)`
Key Variables
The most ordered from aisle is fresh vegeatables

#Plot
```{r}
instacart %>%
  group_by(aisle) %>%
  summarize(n = n()) %>% 
  filter(n>10000) %>% 
  arrange(desc(n)) %>% 
  ggplot(aes(x = reorder(aisle, n), y = n)) + 
    geom_point() + geom_line() + 
    theme(axis.text.x = element_text(size=6, angle = 45, hjust = 1))
  
```

#Table showing the tree most popular items in each of the aisles: baking ingredients, dog food care, packaged vegetables fruits
```{r}
instacart %>%
  filter(aisle %in% c('baking ingredients', 'dog food care', 'packaged vegetables fruits')) %>% 
  group_by(aisle, product_name) %>% 
  summarize(n = n()) %>%  
  arrange(aisle, desc(n)) %>%
  slice(1:3)
  
```
#Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table)

```{r}
instacart %>%
  filter(product_name %in% c('Pink Lady Apples', 'Coffee Ice Cream')) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_time = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = product_name,
    values_from = mean_time)  

  #make 2 by 7.. format days
```

##Problem 2
```{r}
library(p8105.datasets)
data("instacart")
data("BRFSS")

brfss <- 
    brfss_smart2010 %>% 
    janitor::clean_names() %>%
  filter(topic=="Overall Health") %>% 
  mutate(response = as.factor(response)) %>% 
  mutate(response =
           factor(response,levels(response)[c(4,2,3,5,1)]))


brfss %>% 
  filter(year==2002) %>%
  group_by(locationabbr) %>% 
  summarize(n_state =n_distinct(locationdesc)) %>% 
  filter(n_state>=7)

```
In 2002, Connectiticut, Florida, Massachusetts, North Carolina, New Jersey, Pennsylvania each were observed at 7 or more locations
```{r}
brfss %>% 
  filter(year==2010) %>%
  group_by(locationabbr) %>% 
  summarize(n_state =n_distinct(locationdesc)) %>% 
  filter(n_state>=7)
```
In 2010, CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH were observed at 7 or more locations

#Limited to excellent responses
```{r}
spaghetti<-brfss %>% 
  filter(response=="Excellent") %>% 
  group_by(year, locationabbr) %>% 
  summarize(mean_data_value=mean(data_value))
```
#Make a spaghetti plot
```{r}
spaghetti %>% ggplot(aes(x = year, y = mean_data_value, color = locationabbr)) +
  geom_line(aes(group = locationabbr)) +
  geom_point()+ 
  viridis::scale_fill_viridis(discrete = TRUE)
```
```{r}
brfss %>%
  filter(year == c(2006, 2010), locationabbr=="NY") %>%
  ggplot(aes(x = response, y = data_value))+
  geom_violin(aes(fill=response), alpha = .4)+
  facet_grid(~year)+ 
  theme(legend.position = "bottom", axis.text.x = element_text(size=5))

```






## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
