---
title: "p8105_hw3_dd2948"
author: "David DeStephano"
date: "October 7, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(tidyverse)
```

##Problem 1
```{r}
library(p8105.datasets)
data("instacart")

instacart %>%
  group_by(aisle) %>%
  summarize(n = n()) %>% 
  arrange(desc(n))

#Alternatively:
count(instacart, aisle)
#Fresh vegatables is the most popular

```

The dimensions of this data set are `r dim(instacart)`
Key Variables
The most ordered from aisle is fresh vegeatables

#Plot
```{r}
instacart %>%
  group_by(aisle) %>%
  summarize(n = n()) %>% 
  filter(n>10000) %>% 
  arrange(desc(n)) %>% 
  ggplot(aes(x = reorder(aisle, n), y = n)) + 
    geom_point() + geom_line() + 
    theme(axis.text.x = element_text(size=6, angle = 45, hjust = 1))
  
```

#Table showing the tree most popular items in each of the aisles: baking ingredients, dog food care, packaged vegetables fruits
```{r}
instacart %>%
  filter(aisle %in% c('baking ingredients', 'dog food care', 'packaged vegetables fruits')) %>% 
  group_by(aisle, product_name) %>% 
  summarize(n = n()) %>%  
  arrange(aisle, desc(n)) %>%
  slice(1:3)
  
```
#Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table)

```{r}
instacart %>%
  filter(product_name %in% c('Pink Lady Apples', 'Coffee Ice Cream')) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_time = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = product_name,
    values_from = mean_time)  

  #make 2 by 7.. format days
```

##Problem 2
```{r}
library(p8105.datasets)
data("instacart")
data("BRFSS")

brfss <- 
    brfss_smart2010 %>% 
    janitor::clean_names() %>%
  filter(topic=="Overall Health") %>% 
  mutate(response = as.factor(response)) %>% 
  mutate(response =
           factor(response,levels(response)[c(4,2,3,5,1)]))


brfss %>% 
  filter(year==2002) %>%
  group_by(locationabbr) %>% 
  summarize(n_state =n_distinct(locationdesc)) %>% 
  filter(n_state>=7)

```
In 2002, Connectiticut, Florida, Massachusetts, North Carolina, New Jersey, Pennsylvania each were observed at 7 or more locations
```{r}
brfss %>% 
  filter(year==2010) %>%
  group_by(locationabbr) %>% 
  summarize(n_state =n_distinct(locationdesc)) %>% 
  filter(n_state>=7)
```
In 2010, CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH were observed at 7 or more locations

#Limited to excellent responses
```{r}
spaghetti<-brfss %>% 
  filter(response=="Excellent") %>% 
  group_by(year, locationabbr) %>% 
  summarize(mean_data_value=mean(data_value))
```
#Make a spaghetti plot
```{r}
spaghetti %>% ggplot(aes(x = year, y = mean_data_value, color = locationabbr)) +
  geom_line(aes(group = locationabbr)) +
  geom_point()+ 
  viridis::scale_fill_viridis(discrete = TRUE)
```
```{r}
brfss %>%
  filter(year == c(2006, 2010), locationabbr=="NY") %>%
  ggplot(aes(x = response, y = data_value))+
  geom_violin(aes(fill=response), alpha = .4)+
  facet_grid(~year)+ 
  theme(legend.position = "bottom", axis.text.x = element_text(size=5))

```

##Problem 3
```{r}
accel<-read_csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "activity", 
    values_to = "activity_count") %>% 
  mutate(weekend=ifelse(day==("Saturday") | day==("Sunday"), 1, 0 )) %>% 
mutate(day=as.factor(day),
       day =
           factor(day,levels(day)[c(2,6,7,5,1, 3,4)]))


```
#reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).
I am not sure what why the variable classes would not be reasonable
The data set includes 6 variables: week, day_id, day, activity, and weekend. There are 50400 rows and 6 columns, dimensions: `r dim(accel)`

#Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?
The sum for each specific day using the day id:
```{r}
accel %>% 
  group_by(day_id) %>% 
  summarise(sum(activity_count))
```
The combined sums for each day of the week using the day of the week variable:
```{r}
accel %>% 
  group_by(day) %>% 
  summarise(sum(activity_count))
```
No trends seem to be apparent. Friday has the highest daily activity levels, while saturday has the lowest, so maybe this person is very active on fridays but then sits around on saturdays. He seems to become most active from the middle of the week until friday

#Ignore this plot
```{r}
accel %>% ggplot(aes(x=activity, y=activity_count, color = day)) +
  geom_line()+
  geom_point()

```


#Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week.

```{r}
#Create an hour variable
accel2<- accel %>%
  mutate(activity2=sapply(strsplit(activity, split='_', fixed=TRUE), function(x) (x[2]))) %>% 
  mutate(hour=as.numeric(activity2)/60) %>% 
  mutate(hour2=ifelse(hour<=1, 1, 
ifelse(1<hour&hour<=2, 2, 
ifelse(2<hour&hour<=3, 3, 
       ifelse(3<hour&hour<=4, 4,
              ifelse(4<hour&hour<=5, 5,
                     ifelse(5<hour&hour<=6, 6,
                            ifelse(6<hour&hour<=7, 7,
                                   ifelse(7<hour&hour<=8, 8,
                                          ifelse(8<hour&hour<=9, 9,
                                                 ifelse(9<hour&hour<=10, 10,
                                                        ifelse(10<hour&hour<=11, 11,
                                                               ifelse(11<hour&hour<=12, 12,
                                                                      ifelse(12<hour&hour<=13, 13,
                                                                             ifelse(13<hour&hour<=14, 14,
                                                                                    ifelse(14<hour&hour<=15, 15,
                                                                                           ifelse(15<hour&hour<=16, 16,
                                                                                                  ifelse(16<hour&hour<=17, 17,
                                                                                                         ifelse(17<hour&hour<=18, 18,
                                                                                                                ifelse(18<hour&hour<=19, 19,
                                                                                                                       ifelse(19<hour&hour<=20, 20,
                                                                                                                              ifelse(20<hour&hour<=21, 21,
                                                                                                                                     ifelse(21<hour&hour<=22, 22,
                                                                                                                                            ifelse(22<hour&hour<=23, 23,
                                                                                                                                                   ifelse(23<hour&hour<=24, 24,0)))))))))))))))))))))))))
      #Is there a better way to do this with sapply or a for loop?   
         
#Plot
accel2 %>% 
  group_by(hour2, day) %>% 
  summarize(activity_hour=mean(activity_count)) %>% 
  ggplot(aes(x=hour2, y=activity_hour, color=day))+
  geom_point()+
  geom_line()


```   
The trend is that in the morning (12 AM to 4) this person is most likely asleep every day (except on thursday when he seems to sometimes be awake). He then wakes up and has a more or less constant activity level throughout the dat(except sunday) until 6PM, where he begins to either wind down on certain days, or exercise on other days.

 


#If you want each day instead of hour for each day you can run the following code:
```{r}
accel %>% 
  group_by(day_id, day) %>% 
  summarize(activity2=sum(activity_count)) %>% 
  ggplot(aes(x=day_id, y=activity2, color=day))+
  geom_point()+
  geom_line()
```



```{r}
accel2 %>% 
  group_by(hour2, day, day_id) %>% 
  summarize(activity_hour=mean(activity_count)) %>% 
  ggplot(aes(x=hour2, y=activity_hour, color=day))+
  geom_point()+
  geom_line()
```







